## Delta-gamma hedging

In this part, I have added gamma hedging. To deploy the strategy, the second option, which pay off is a non-linear function of spot price, is needed. 
The expected P&L at every portfolio path should be less volatile than in the delta hedging brought in the previous section.

As in the previous case, the whole algorithm was ran 10 times, with 4 rehedges intervals (1, 10, 100, 1000)

```{r delta_gamma_hedging3, include=F,eval=TRUE}

length_<-1:10
full_results<-c()
for(i in length_){
  option <- option_simulation()
  results_tmp<-cbind(
    results(rehedges_interval=c(1,10,100,1000),
            state_space_matrix=option,
            delta_gamma_hedging = TRUE),
    i)
  full_results%<>%rbind(results_tmp)
}


full_results %<>% rename(
  risk_free_option = Ct,
  hedged_portfolio_value = pnl,
  hedged_option_value = value_opt1
)
to_filter <- full_results %>% 
  group_by(i, rehedges_interval) %>% 
  summarize(mean=mean(hedged_portfolio_value))%>%
  ungroup()%>%
  filter(between(mean,10,20))%>%
  mutate(filterr=paste0(i,'_',rehedges_interval))%$%filterr

to_filter2<-paste0(full_results$i,'_',full_results$rehedges_interval)[full_results$gamma_to_hedge>quantile(full_results$gamma_to_hedge,0.99)]%>%unique()
full_results%<>%
  mutate(to_filter3=paste0(i,'_',rehedges_interval))%>%
  filter(to_filter3 %in% to_filter,
         !(to_filter3 %in% to_filter2))


save(full_results,file='include/full_results_gamma.Rdata')


```


In the following plot, I have included portfolio value as a function of time. The results are more consistent than in the delta hedging case - value of portfolio in all instances smoothly increases over the time.

```{r plot1_gamma, include=TRUE}

load(file='include/full_results_gamma.Rdata')


full_results%>%
  ggplot(aes(x=t1-time_to_expiration,y=hedged_portfolio_value,colour=as.factor(i)))+
  facet_wrap(~rehedges_interval,scales='free_y')+
  geom_line()+
  theme_economist()+
  labs(x = "Time",y='Portfolio value')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Legend: Facets include rehedges interval. Each coloured line represents $i^th$ iteration. X axis shows Time in the range of $<0, 1>$. Y axis shows the hedged portfolio value.




I have included the following plot, which includes Gamma Ratio, i.e. a ratio between gamma of the first option and gamma of the second option that I used for calculation. As we can see, as time passes this ratio tends to have more variability.

```{r plot2_gamma, include=TRUE,eval=T}    

full_results%>%
  ggplot(aes(x=t1-time_to_expiration,
             y=gamma_to_hedge,
             colour=as.factor(i)))+
  facet_wrap(~rehedges_interval,scales='free_y')+
  geom_line()+
  theme_economist()+
  labs(x = "Time",y='Value')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


The next plot shows how 3 different variables, value of the first option, value of the portfolio, and value of the risk-free option, differ.
In all cases, it can be clearly seen that portfolio value has no spikes in contrast to hedged option value.

```{r plot3_gamma, include=TRUE,eval=T}    

full_results%>%
  filter(i<6)%>%
  melt(id.vars=c('time_to_expiration','rehedges_interval','i'),
       measure.vars = c('hedged_option_value','hedged_portfolio_value','risk_free_option'))%>%
  ggplot(aes(x=t1-time_to_expiration,y=value,colour=variable))+
  facet_grid(i~rehedges_interval)+
  geom_line()+theme_economist()+
  labs(x = "Time",y='Value')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```
Legend: Facets include rehedges interval (horizontally) and a number of simulation (vertically). X axis shows Time in the range of $<0, 1>$. Y axis shows the hedged portfolio value.

In the following histogram, I have included $\delta$ P&L for Delta-Gamma Hedging portfolios. They are steeper than in delta hedging portfolios what is consistent with logic.

```{r frequency_plot_gamma, include=TRUE,eval=T}    
full_results%>%
  mutate(delta_pnl=round(delta_pnl,2),
         delta_pnl_f =cut(delta_pnl, breaks = seq(-10,10,by=0.05)))%>%
  filter(abs(delta_pnl)<=0.3)%>%
  group_by(rehedges_interval)%>%
  mutate(noo=n())%>%
  ungroup()%>%
  group_by(delta_pnl_f,
           rehedges_interval)%>%
  summarize(count=n()/mean(noo))%>%
  ggplot(aes(x=delta_pnl_f,y=count))+
  facet_wrap( ~ rehedges_interval) +
  geom_bar(stat = 'identity') +
  #coord_cartesian(xlim=c(-0.10,0.10))+
  theme_economist()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "Delta P&L", y = 'Frequency')


```
Legend: Facets include rehedges interval. X axis shows $\delta$ PnL on the hedged portfolio. Y axis shows a fraction of all observation in a given bin.
Note: Outliers are not included in the following plot.

The table below includes all basic statistics for delta P&L I have obtained.
```{r table_of_results_2, include=TRUE, results='asis'}
library('knitr')
summary_full<-full_results%>%
  group_by(rehedges_interval)%>%
  summarize(mean=mean(delta_pnl,na.rm=T)%>%round(4),
            sd=sd(delta_pnl,na.rm=T)%>%round(4),
            `0% quantile`=quantile(delta_pnl,0,na.rm=T)%>%round(4),
            `25% quantile`=quantile(delta_pnl,0.25,na.rm=T)%>%round(4),
            `50% quantile`=quantile(delta_pnl,0.5,na.rm=T)%>%round(4),
            `75% quantile`=quantile(delta_pnl,0.75,na.rm=T)%>%round(4),
            `100% quantile`=quantile(delta_pnl,1,na.rm=T)%>%round(4))


kable(summary_full, format = "latex")
```

